---
title: "Project_Fast&Serious"
output: html_document
---
#Install required packages
```{r}
require(tidyverse)
require(ggplot2)
library(dplyr)
library(lubridate)
library(ggplot2)
```

#Load the data
```{r}
Covid_19 <-read_csv('Final_Covid19_Dataset.csv')

Covid_19
```

#Data Cleaning 
#Convert date into 'date' format
```{r setup, include=FALSE}
class(Covid_19$Date)

Covid_19$Date <- as.Date(Covid_19$Date, "%m/%d/%Y")

sapply(Covid_19, class)
Covid_19
```

#How is the number of cases increasing per day globally?
```{r}
#Total no of increasing casees per day globally
globally_cases <- Covid_19%>% group_by(Date)%>%
              summarise(No_of_cases_increased_per_day=sum(Confirmed))%>%
              arrange(Date)
globally_cases

globally_cases%>%mutate_all(funs(. - lag(.))) %>% na.omit() 

#First 10 records for analysis in detail 
G_10<-globally_cases%>%head(10)
G_10

#Visualization on initial first 10 records
Cases_per_day <- ggplot(G_10, aes(x=Date, y=No_of_cases_increased_per_day)) + 
    geom_line(aes(group=1), colour="#000099")+
    geom_point(size=3, colour="#CC0000")+
   labs(x= "Date", y="No of cases", title="No of cases rise in one day") 
   

Cases_per_day
```


#Question -> Is there any correlation between a country’s death reported and its healthcare system rankings to provide positive health outcomes and to understand how they’re managing the virus.
```{r}
Heathcare <- Covid_19%>%
            group_by(Country, Healthcare_System_Rank)%>%
            summarise(Total_Deaths = max(Deaths))%>%
           arrange(Healthcare_System_Rank)
Heathcare
```

#Question ->Which countries are most affected by developed or developing countries?
```{r}
Distribution_class <- Covid_19%>%group_by(Classification)%>%
                      summarise(Total_no_of_confirmed_cases = max(Confirmed))

Distribution_class
```


#Question -> Which is the most affected developed country by Covid-19 outbreak?
```{r}
Max_count_developed <- Covid_19%>%group_by(Classification,Country)%>%
                      summarise(Total_no_of_confirmed_cases = max(Confirmed))%>%
                      arrange(desc(Total_no_of_confirmed_cases))

Max_count_developed
```


#Question -> Which country has the highest fatality rate?
```{r}
F_rate_all<- Covid_19 %>% filter(Date == "2020-04-07")%>%mutate(Fatality_Rate = (Deaths/Confirmed)*100)

F_rate <- F_rate_all%>%select(Country,Recovered,Confirmed,Fatality_Rate)%>%arrange(desc(Fatality_Rate))
F_rate

#Top 10 high Fatality rate
F_Top_10<-F_rate %>%head(15)
F_Top_10
#Data visualization on top 10 high Fatality rate countries

F_visual <- ggplot(data=F_Top_10, aes(x=reorder(Country, -Fatality_Rate), y=Fatality_Rate)) + geom_col(width=0.5, fill='#E69F00',color="darkred")+theme(axis.text.x = element_text(angle = 60, hjust = 1))+labs(x= "List of Countries", y="Fatality rate (%)",title="Fatality rate by Countries")

F_visual

```

#Question -> Which country has the highest mortality rate?
```{r}
M_rate_all<- Covid_19 %>% filter(Date == "2020-04-07")%>%mutate(Mortality_Rate = (Recovered/Confirmed)*100)

M_rate <- M_rate_all%>%select(Country,Recovered,Confirmed,Mortality_Rate)%>%arrange(desc(Mortality_Rate))

#Top 10 high mortality rate
M_Top_10<-M_rate %>%head(10)

#Data visualization on top 10 high mortality rate countries
M_visual<-M_Top_10%>%
  ggplot( aes(x=reorder(Country, Mortality_Rate), y=Mortality_Rate)) +
    geom_bar(stat="identity", fill="#FF9999", alpha=.6, width=.4,  colour="black")+
    coord_flip() +
    labs(x= "List of Countries", 
             y="Mortality rate(%)", 
             title="Mortality rate by Countries") +
    theme_bw()

M_visual

```

#Question -> Which country has the highest number of active cases? #Filter data for 4/7/2020
```{r}
Active_cases <- Covid_19%>%group_by(Country)%>%filter(Date == "2020-04-07")%>%  summarise(ConfirmedCases=max(Confirmed), 
            RecoveredCases = max(Recovered), DeathsCases = max(Deaths)) %>%    
            mutate(Active_cases = ConfirmedCases-(RecoveredCases+DeathsCases))%>%
            arrange(desc(Active_cases))
Active_cases

```

#Question -> Countries with most affected cases based upon count per 100k people?

```{r}
latest_date_data=Covid_19%>%filter(Date=="2020-04-07")

count_per_100k=latest_date_data%>%mutate(Confirmed_per_100k=as.numeric(format(round((100000*Confirmed)/Total_Population),2)))
count_per_100k

count_per_100k_descending = count_per_100k[order(round(as.numeric(count_per_100k$Confirmed_per_100k)),decreasing = TRUE, na.last = TRUE), ]
count_per_100k_descending%>%top_n(n=10, wt=Confirmed_per_100k)%>%ggplot(aes(x=reorder(Country, Confirmed_per_100k), y=Confirmed_per_100k, fill=Country)) + geom_bar(stat="identity")+geom_text(aes(label = Confirmed_per_100k))+ coord_flip()+labs(x="Country", y="Confirmed_per_100k", title = "Top 10 affected countries based on Count_per_100k")
```

#Question -> How COVID-19 affecting people from a different age group?

```{r}
Median_age_data <- latest_date_data$Median_age
latest_date_data$Median_Age_Group <- cut(Median_age_data, 4)
latest_date_data%>%summarise(sum(Confirmed))

latest_date_data%>%group_by(Median_Age_Group)%>%summarise(sum(Confirmed),num_of_countries=n())
```

#Question -> What is the correlation between confirmed cases and Classification ?

```{r}
ggplot(latest_date_data, aes(x = Total_Population, y = Confirmed , color = Classification)) + scale_y_log10() + scale_x_log10()+ geom_point()
```


#Question: Which country has been most affected even though they have good healthcare facilities?

```{r}
death_with_health_care <- latest_date_data %>% select(Country,Deaths,Healthcare_System_Rank) %>% arrange(Healthcare_System_Rank)%>% head(10)
death_with_health_care
health_care_graph <- ggplot(death_with_health_care, aes(x = Healthcare_System_Rank, y = Deaths, color = Country)) + geom_point(size=3)
health_care_graph

```

